Details of specific functions

### 2020-09-11 inst/gitbook/03_specific-tutorials.Rmd
Added manual and test!
Nearly there!

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`ape`](https://cran.r-project.org/web/packages/vegan/index.html) (but see also [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree`) |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

You can find detailed explanation on how between group metrics work [here](#betweengroupmetricsexplain).
The `group.dist` metric allows to measure the distance between two groups in the multidimensional space.
This function needs to intake several groups and use the option `between.groups = TRUE` in the `dispRity` function.
It calculates the vector normal distance (euclidean) between two groups and returns 0 if that distance is negative.
Note that it is possible to set up which quantiles to consider for calculating the distances between groups.
For example, one might be interested in only considering the 95% CI for each group.
This can be done through the option `probs = c(0.025, 0.975)` that is passed to the `quantile` function.
It is also possible to use this function to measure the distance between the groups centroids by calculating the 50% quantile (`probs = c(0.5)`).

@@

@@

```{r}
best.metric <- function() return(42)
```
Joke aside, this is a legitimate question that has no simple answer: **it depends on the dataset and question at hand**.
Thoughts on which metric to choose can be find in @moms and @Guillerme2020 but again, will ultimately depend on the question and dataset.
The question should help figuring out which type of metric is desired: for example, in the question "does the extinction released niches for mammals to evolve", the metric in interest should probably pick up a change in size in the trait space (the release could result in some expansion of the mammalian morphospace); or if the question is "does group X compete with group Y", maybe the metric of interested should pick up changes in position (group X can be displaced by group Y).

>>>>

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`ape`](https://cran.r-project.org/web/packages/vegan/index.html) (but see also [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree`) |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

You can find detailed explanation on how between group metrics work [here](#betweengroupmetricsexplain).

##### `group.dist`

The `group.dist` metric allows to measure the distance between two groups in the multidimensional space.
This function needs to intake several groups and use the option `between.groups = TRUE` in the `dispRity` function.
It calculates the vector normal distance (euclidean) between two groups and returns 0 if that distance is negative.
Note that it is possible to set up which quantiles to consider for calculating the distances between groups.
For example, one might be interested in only considering the 95% CI for each group.
This can be done through the option `probs = c(0.025, 0.975)` that is passed to the `quantile` function.
It is also possible to use this function to measure the distance between the groups centroids by calculating the 50% quantile (`probs = c(0.5)`).

@@

##### `point.dist`

The metric measures the distance between the elements in one group (`matrix`) and a point calculated from a second group (`matrix2`).
By default this point is the centroid but can be any point defined by a function passed to the `point` argument.
For example, the centroid of `matrix2` is the mean of each column of that matrix so `point = colMeans` (default).
This function also takes the `method` argument like previous one described above to measure either the `"euclidean"` (default) or the `"manhattan"` distances:

```{r}
## Measuring the distance between the elements of the first group
## and the centroid of the second group
summary(dispRity(grouped_space, metric = point.dist,
                 between.groups = TRUE))

## Measuring the distance between the elements of the second group
## and the centroid of the first group
summary(dispRity(grouped_space, metric = point.dist,
                 between.groups = list(c(2,1)))))

## Measuring the distance between the elements of the first group
## a point defined as the standard deviation of each column
## in the second group
sd.point <- function(matrix2) {apply(matrix2, 2, sd)}
summary(dispRity(grouped_space, metric = point.dist,
                 point = sd.point, method = "manhattan",
                 between.groups = TRUE))
```

@@

```{r}
best.metric <- function() return(42)
```

Joke aside, this is a legitimate question that has no simple answer: **it depends on the dataset and question at hand**.
Thoughts on which metric to choose can be find in @moms and @Guillerme2020 but again, will ultimately depend on the question and dataset.
The question should help figuring out which type of metric is desired: for example, in the question "does the extinction released niches for mammals to evolve", the metric in interest should probably pick up a change in size in the trait space (the release could result in some expansion of the mammalian morphospace); or if the question is "does group X compete with group Y", maybe the metric of interested should pick up changes in position (group X can be displaced by group Y).
